Optimal Algorithms for Dynamic Balanced Graph Partitioning

### 2020-05-14 paper/disc20clustering.tex
wip

	Next, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
	Fix a~partition for \OPT prior serving the sequence $\sigma_E$.
	Even without migrations, \OPT may mitigate paying for requests between at most $3\ell$ pairs of nodes that can be collocated in its clusters.
	Recall that $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), and the counter of an unsaturated pair is at most $\alpha - 1$.
	Hence, $\OPT$ may mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
	Finally, during the phase, \OPT is faced with at least $\chi := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests from $\sigma_E$ that it cannot mitigate.

	By migrating nodes, \OPT may decrease its cost, and pay less than $\chi$ for requests from $\sigma_E$.
	By swapping a~pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
	As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
	By performing a~single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
	Hence, for serving at least $\chi$ unavoidable requests from $\sigma_E$, $\OPT$ pays at least
	%
	\[
		\OPT(\sigma_E) + \OPT(M) \geq \chi \cdot \frac{2\alpha}{4 (\alpha-1)}\geq \frac{|\sigma_E|} {2} - 2 \alpha \cdot \ell.
	\]
	%	
	To bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	For succintness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
%
	\[
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
	\]
%	
	\medskip

>>>>

	Next, we bound $\OPT(\sigma_E) + \OPT(M)$.
	The sequence $\sigma_E$ accounts only for unsaturated edges, thus there are at most $\alpha-1$ requests to each pair in $\sigma_E$.
	\OPT may have at most $3\ell$ pairs of nodes collocated in its clusters, and thus avoid paying for $3\ell\cdot(\alpha-1)$ requests from $\sigma_E$.
	Hence, at least $\chi := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests from $\sigma_E$ are external requests with respect to \OPT's configuration at the beginning of the phase.
	Faced with these requests, \OPT may serve them remotely or perform migrations to decrease its cost.
	By swapping a~pair of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
	This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterward.
	Hence, by performing a~single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
	The total cost of \OPT is then at least
	%
	\[
		\OPT(\sigma_E) + \OPT(M) \geq \chi \cdot \frac{2\alpha}{4 (\alpha-1)}\geq \frac{|\sigma_E|} {2} - 2 \alpha \cdot \ell.
	\]

	Finally, to bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	For succintness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
%
	\[
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
	\]
%	
	\medskip
